### 2016-09-21 mcmc.tex
skilling
frenchspacing
Added comments

\newcommand{\MCMC}{\acronym{MCMC}}
\newcommand{\MH}{\acronym{M--H}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\data}{D}
\newcommand{\pars}{\theta}
\newcommand{\best}{{\mathrm{(best)}}}
\newcommand{\better}{{\mathrm{(better)}}}

@@

@@

The simplest algorithm for \MCMC\ is the Metropolis--Hastings algorithm (\MH\ \MCMC).\note{%
  There are many claims that this algorithm is incorrectly named,
  with claims that it should be credited to Enrico Fermi or Stan Ulam.
  We don't have any opinions of this matter, but encourage the reader to follow this up.
  The original paper is \citet{metropolis}, and there are a few sketchy historical notes
  in \citet{geyer}.}
It is so simple that we recommend that any reader of this document
  who has not previously implemented the algorithm take a break at the end
  of this \sectionname\ and implement it forthwith, in a short piece of computer code,
  in the context of some simple problems.\note{%
    We request this in a trivial case below in
    \problemname~\ref{prob:MH} and the following problems in this
    \sectionname.
    We make the same request in a less trivial context in our
    model-fitting \documentname\ (\citealt{fittingaline},
    \problemname~6 and 7), where we ask the Reader to implement \MCMC\ for
    a useful mixture model.}

@@

HOGG SAY HERE: ...You must describe your priors in the same parameterization that you ``step'' in.
That is, if you change your parameters, you \emph{must} change your priors correspondingly.
Give the example of going from $A$, $\phi$ to $A\,\cos\phi$, $A\,\sin\phi$.

@@

\begin{problem}\label{prob:itertau}
Write a peice of code that estiamtes the integrated autocorrelation time for a
chain of samples using an estimate of the autocorrelation function and a given
``window'' size $M$ (see \equationname~\ref{eq:tauest}).
Plot the estimated $\tau$ as a function of $M$ for several contiguous segments
of the chain and overplot the sample function based on the full chain.
What can you conclude from this plot?
Implement an iterative procedure for automatically choosing $M$.\note{%
The recipe given on page 16 of A.~Sokal's notes
(\url{http://www.stat.unc.edu/faculty/cji/Sokal.pdf}) might be helpful.
Note that the definition of $\tau$ in that document is half of the definition
that we use.}
Overplot this estimate on the plot of $\tau(M)$ and the result should look
like \figurename~\ref{fig:itertau}.
\end{problem}

@@

\paragraph{Gibbs:}
In many problems, the parameters---the components or entries of the
blob $\pars$---are not equivalent at all.
Some are global parameters, which touch or have an effect on every
sub-part of the data blob $\data$, while others are local parameteres,
which only touch a small subset of the data.
Or, in another use case, some parameters are linear parameters
that---at fixed values of the other parameters---will have a Gaussian
likelihood or posterior pdf.
In these problems, there are some directions in parameter space that
are hard to move in (they require complete re-calculation of the full
non-linear function $f(\pars)$, and other directions that are very
easy to move in (they require only partial re-calculations or the
sampling could even be analytic and exact).
For problems with these structures, Gibbs samplers are ideal, and
there are good implementations.\note{For example, \project{JAGS}; find it on the internets.}

@@

  Brewer,~B.~J., P\'artay,~L.~B. & Cs\'anyi,~G., 2011,

@@

>>>>

\newcommand{\MCMC}{\acronym{MCMC}}
\newcommand{\MH}{\acronym{M--H}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\data}{D}
\newcommand{\pars}{\theta}
\newcommand{\bl}[1]{{\color{red}{[BL:] #1}}}
\newcommand{\best}{{\mathrm{(best)}}}
\newcommand{\better}{{\mathrm{(better)}}}

@@

\bl{
Overall comments:
\begin{itemize}
\item Would it be useful to mention that MCMC is useful when you cannot or don't want to evaluate the posterior on a grid?
\item $p(\theta)$ is used in place of $p(\theta|D)$ in various places. This is mildly perturbing - I would either change it to $p(\theta|D)$ everywhere or explicitly say that the shorter convention $p(\theta)$ is used.
\item How much do you want to go into non Metropolis-H algorithms? At the moment this is either not sufficiently or too detailed. I would either reduce it to a brief enumeration or to a discussion of features, advantages and drawbacks of various methods. I am happy to help with that. However, it might be hard to avoid discussing specific technologies and codes. Worse: I am not convinced by the practical usefulness of a generic summary of all sampling methods.
\end{itemize}
}

@@

The simplest algorithm for \MCMC\ is the Metropolis--Hastings algorithm (\MH\ \MCMC).\note{%
  There are many claims that this algorithm is incorrectly named,
  with claims that it should be credited to Enrico Fermi or Stan Ulam.
  We don't have any opinions of this matter, but encourage the reader to follow this up.
  The original paper is \citet{metropolis}, and there are a few sketchy historical notes
  in \citet{geyer}.}
It is so simple that we recommend that any reader of this document
  who has not previously implemented the algorithm takes a break at the end
  of this \sectionname\ and implement it forthwith, in a short piece of computer code,
  in the context of some simple problems.\note{%
    We request this in a trivial case below in
    \problemname~\ref{prob:MH} and the following problems in this
    \sectionname.
    We make the same request in a less trivial context in our
    model-fitting \documentname\ (\citealt{fittingaline},
    \problemname~6 and 7), where we ask the Reader to implement \MCMC\ for
    a useful mixture model.}

@@

HOGG SAY HERE: ...You must describe your priors in the same parameterization that you ``step'' in.
That is, if you change your parameters, you \emph{must} change your priors correspondingly.
Give the example of going from $A$, $\phi$ to $A\,\cos\phi$, $A\,\sin\phi$. \bl{Isn't that discussed somewhere below?}

@@

\begin{problem}\label{prob:itertau}
Write a piece of code that estimates the integrated autocorrelation time for a
chain of samples using an estimate of the autocorrelation function and a given
``window'' size $M$ (see \equationname~\ref{eq:tauest}).
Plot the estimated $\tau$ as a function of $M$ for several contiguous segments
of the chain and overplot the sample function based on the full chain.
What can you conclude from this plot?
Implement an iterative procedure for automatically choosing $M$.\note{%
The recipe given on page 16 of A.~Sokal's notes
(\url{http://www.stat.unc.edu/faculty/cji/Sokal.pdf}) might be helpful.
Note that the definition of $\tau$ in that document is half of the definition
that we use.}
Overplot this estimate on the plot of $\tau(M)$ and the result should look
like \figurename~\ref{fig:itertau}.
\end{problem}

@@

\paragraph{Gibbs:}
In many problems, the parameters---the components or entries of the
blob $\pars$---are not equivalent at all.
Some are global parameters, which touch or have an effect on every
sub-part of the data blob $\data$, while others are local parameters,
which only touch a small subset of the data.
Or, in another use case, some parameters are linear parameters
that---at fixed values of the other parameters---will have a Gaussian
likelihood or posterior pdf.
In these problems, there are some directions in parameter space that
are hard to move in (they require complete re-calculation of the full
non-linear function $f(\pars)$, and other directions that are very
easy to move in (they require only partial re-calculations or the
sampling could even be analytic and exact).
For problems with these structures, Gibbs samplers are ideal, and
there are good implementations.\note{For example, \project{JAGS}; find it on the internets.}

@@

  Brewer,~B.~J., P\'artay,~L.~B. \& Cs\'anyi,~G., 2011,

@@

\bibitem[Skilling(2006)]{skilling}
  Skilling, J., 2006,
  Nested sampling for general bayesian computation,
  \textit{Bayes.\,Anal.}\ \textbf{1}(4) 833-–860
  \doi{10.1214/06-BA127}.
